---
title: "Build Online Apps with R"
subtitle: "Data"
author: "Matthias Bannert"
date: "30 Nov 2018"
output: ioslides_presentation
widescreen: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Data Generating Processes (DGP)

- research (field / lab / surveys)
- simulation
- measuring
- tracking
- ... 

## Types of Datasets: Time Series

```{r}
ts2 <- ts(rnorm(20),
          start = c(1995,1),
          frequency = 4)
ts2
```

examples: monthly revenues over time, stocks,
aggregated log files



## Types of Datasets: Cross Sectional Data

```{r}
head(mtcars)
```

- multiple variables
- one period



## Types of Datasets: Panel Data

- multiple variables
- longitudinal data 
- e.g. German Socio-Economic Panel (GSOEP)

## Types of Datasets: Nested Data

```{r}
l <- list()
l$element1 <- 2
l$element2 <- head(mtcars,4)
l
```

examples: meta information, sector classification (hierarchical), GDP components, translations, attributes, properties, [Geo spatial data](https://github.com/mbannert/maps/blob/master/ch_bfs_regions.geojson)

## Task I: Dataset types

Represent each of the following dataset types using R: 

- time series
- cross section
- panel
- a nested structure.

Discuss suggest an in-memory representation (inside the R session) and a file based representation. 

## Data Management

**storage**

- file based
- database
- relational
- non-relational
- hybrid solutions
- centralized / decentralized 

**granularity**

- micro data / on-the-fly computation (e.g. OLAP cubes)
- aggregated data 

## An Exploring Datasets Strategy: Reading Data

Read data into R
- readr package
- foreign package
- openxlsx

## An Exploring Datasets Strategy: Checking Data    

Check whether it worked
- head()
- tail()
- str()
- ls()

## An Exploring Datasets Strategy: Get to Know the Data    

Get and idea of the data
- summary
- distribution
- missing values
- scatter plots

## Task II: Scaling it up

1. Install the *fivethirtyeight* package

2. pick your favorite dataset

3. explore the dataset, report your findings (summary stats, visual)

4. How can we do this for multiple datasets? Write a function to get a summary statistics for every
dataset in fivethirtyeight... 


## Data Sources

- official statistics ([Live Stream on Election Day](https://opendata.swiss/de/dataset/echtzeitdaten-am-abstimmungstag))
- universities
- journals
- open data repositories ([nature on opendata](- https://www.nature.com/news/peer-review-activists-push-psychology-journals-towards-open-data-1.21549), [presentation on opendata in R by Thomas Lo Russo](https://openzh.github.io/opendata_and_r))
- commercial providers, e.g., https://www.kaggle.com/datasets

## Example Sources: Websites

- https://opendata.swiss/de/
- http://data.europa.eu/
- https://opentransportdata.swiss/en/
- http://opendata.cern.ch/
- https://www.apa.org/research/responsible/data-links.aspx

## API example: the KOF data API

```{r}
library(kofdata)
library(tstools)
tsplot(get_time_series("ch.kof.barometer"))
## remember the :: syntax to quickly explore
## functions of a (small) package
## kofdata:: + 'hit tab'
```

## List of API Wrappers

- https://github.com/RomanTsegelskyi/r-api-wrappers
- ... 















