---
title: "Build Online Apps with R"
subtitle: "Warm up"
author: "Matthias Bannert"
date: "30 Nov 2018"
output: ioslides_presentation
widescreen: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Goals

- learn about visualization
- solidify R (programming) skills 
- Understand R-Ecosystem

## Organization of the Course

- group work (2-4 students)
- Plenty of time in between sessions to recap and ask questions, study R 
- Short documentation / course project description term paper (1-2 pages)
- App project can be freely chosen
- **Creating a GUI for any R package is blueprint for app ideas!**

## Exam (Semesterleistung)

**Grading Criteria (Individual Grades)**

- Get an app done
-  Idea of the app
- delta (not absolute gain)
- how you organize collaboration

**HAND IN BEFORE SUBMISSION DEADLINE!**



## You STILL know all of those, don't you?

```{r,eval=F,echo=T}
?sum()
ls()
c()
matrix()
data.frame()
list()
head()
tail()
str()
function()
lapply()
```

=> create your own R cheat sheets!

## R Studio - Moving around

- switch from console to script and back (Ctrl+1,2)
- hashes indicate comments, multi-line comments: ctrl+shit+c
- ctrl + enter run selected code 


## Some simple (?) tasks?

** Questions **

  1. Download the MLB dataset and read it into R.
  2. How many observations are in the dataset?
  3. Who is the tallest / smallest player?
  4. Are players older than 32 significantly heavier than younger players?
  5. Create a function that converts inches to cm











